---
title: "Lesson 2: The Statistical Process & Design of Studies"
author: "Homework"
output:
  html_document:
    css: buttondown.css
    toc: no
  pdf_document:
    toc: no
  word_document:
    toc: no
params:
  key: no
  plinks: no
  keyname: "blank"
  docname: "blank"
md_document:
  variant: markdown_phpextra
---

```{r,results='asis',echo=FALSE}
if(params$key==TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$keyname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$keyname,"pdf",sep="."),")\n")
  cat("<!--")
  cat("\n")
} 
if(params$key!=TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$docname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$docname,"pdf",sep="."),")\n")
  cat("\n")
}
```


```{r,include=FALSE}
library(ggplot2)
library(ggthemes)
library(pander)
library(mosaic)
panderOptions('keep.line.breaks',TRUE)
source("../../scripts/ggQQline.R")
source("../../scripts/normTail.R")
source("../../scripts/221_Interactive_Functions.R")
```

**Instructions:  You are encouraged to collaborate with other students on the homework, but it is important that you do your own work.  Before working with someone else on the assignment, you should attempt each problem on your own.**


1.	Classify each of the following variables as categorical or quantitative.

    a. Number of new automobiles sold at a dealership on a given day

    b. Gender 

    c. Total number of hours slept in the past 24 hours - rounded to the nearest hour

    d. Handedness (which hand you write with) 

    e. Species of trees sold at a nursery 

    f. Weight in carats of an uncut diamond

    g. Brand name for a pair of running shoes 


2. Explain the difference between an observational study and a designed experiment.


3. Briefly explain each of the following types of sampling procedures.

    a. Simple random sample


    b. Stratified random sample 


    c. Systematic random sample


    d. Cluster sample

```{r,include=FALSE}
startingPoint = 16
k = 20

```


4. State health care regulators wanted to assess how satisfied residents of nursing homes are with the care they receive.  In order to obtain a sample, the regulators used a registry that contained all of the state's nursing homes listed in alphabetical order.  They randomly chose a number between 1 and `r k` and got "`r startingPoint`." They selected the $`r startingPoint`^{th}$ nursing home on the list and every $`r k`^{th}$ nursing home thereafter until they reached the end of the list.  Which sampling method did they use?

    a. Simple random sample
    b. Cluster sample
    c. Systematic random sample
    d. Stratified random sample
    
```{r,include=FALSE}
if(params$key==TRUE){

#solutions
  
  Solution01 = data.frame(Part=c(LETTERS[1:7]),Solution=c("Quantitative","Categorical","Quantitative","Categorical","Categorical","Quantitative","Categorical"))

    Solution2 =  data.frame(Part="-",Solution="In a designed experiment, researchers manipulate the conditions that the participants experience. In an observational study researchers observe the responses of the individuals, without controlling the conditions experienced by the individuals.")
    
    Solution3 =  data.frame(Part=LETTERS[1:4],
                            Solution=c(
                              "If there is a list of all the items of the population, take a randomly selected number of them. (Software can be used to randomly select the items)",
                              "Stratified sampling works well when the items are similar within each stratum and tend to differ from one stratum to another. A certain number of items is randomly selected from each of these strata.",
                              "This sample requires taking every $k^{th}$ item in the population, beginning at a random starting point. Systematic sampling works well when the items are in a random, sequential ordering. If the items are not arranged randomly, a systematic sample can miss important parts of the population.",
                              "A cluster sample consists of taking all items in one or more randomly selected clusters, or blocks. When the variation from one block to another is relatively low, compared to the variation within the block, cluster sampling is a reasonable way to get a sample."))
    Solution4 =  data.frame(Part="-",Solution="c. Systematic Random Sample")

} # end params == TRUE.  We use this so it doesn't have to run for non answer key path.
```


```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){
  cat("-->")
   cat("## Solutions\n\n")
      cat("\n\n **Please note that the steps show rounded numbers, but that the final answers to the problems are calculated without rounding.**")
}
```



```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){

    all_solutions = sort(ls(pattern="Solution"))
    key_list = NULL
    for (i in 1:length(all_solutions)){
      temp = get(all_solutions[i])
      temp$Solution = as.character(temp$Solution)
      key_list = rbind(key_list,data.frame(Problem=i,temp))
    }
    
      pander(key_list,split.cell = 80, split.table = Inf,justify = c( 'center', 'left',"left"))
} # end params == TRUE.  We use this so it doesn't have to run for non answer key path.

```




