---
title: 'Lesson 4: Describing Quantitative Data: Center, & Spread'
author: "Preparation"
output:
  html_document:
    css: buttondown.css
  pdf_document: default
  word_document: default
params:
  key: no
  plinks: no
  keyname: "blank"
  docname: "blank"
md_document:
  variant: markdown_phpextra
---

```{r,results='asis',echo=FALSE}
if(params$key==TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$keyname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$keyname,"pdf",sep="."),")\n")
  cat("<!--")
  cat("\n")
} 
if(params$key!=TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$docname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$docname,"pdf",sep="."),")\n")
  cat("\n")
}
```

**Directions: Please fill in Part I as you study the Reading Assignment. Once you finish the reading, complete the questions on Part II. You may use your notes, the key, and the help videos. Be sure to take this completed assignment to your group meeting where you can ask and help answer questions on this assignment.**

```{r,include=FALSE}
library(ggplot2)
library(ggthemes)
library(pander)
library(mosaic)
library(reshape2)
panderOptions('keep.line.breaks',TRUE)
source("../scripts/ggQQline.R")
source("../scripts/normTail.R")
source("../scripts/221_Interactive_Functions.R")
```

## Problems

**Part I:**  Use the information in the reading assignment to complete these questions.

1. Match the following statistical terms with their appropriate definitions.    
    ___ Shape of the distribution   
    ___ Center of the distribution    
    ___ Spread of the distribution    

    a. Average, mean, median, mode
    b. Standard deviation, variance
    c. Symmetric, bell-shaped, skewed right, and skewed left, uniform, unimodal, bimodal, or multimodal

2. Define standard deviation.

3. Before your Group Preparation session, explain "standard deviation" to a roommate, family member, or other person who has not studied statistics.  Describe the experience.  How did you explain this idea?

4. Define sample variance.

5. Provide a definition and/or a brief description of the following terms.

Percentiles -

Quartiles -

Five-number summary - 

Boxplot - 

**Part II:**  

```{r,include=FALSE}

Mystery <- read.csv("https://raw.githubusercontent.com/byuistats/data/master/MysteryInvestmentOpportunities/MysteryInvestmentOpportunities.csv",header = TRUE, stringsAsFactors = FALSE)

problem_6_CompA <- favstats(Mystery$A)
meanA <- round(problem_6_CompA$mean,3)
medianA <- round(problem_6_CompA$median,3)
sdA <- round(problem_6_CompA$sd,3)
sampleA <- problem_6_CompA$n

problem_6_CompB <- favstats(Mystery$B)
meanB <- round(problem_6_CompB$mean,3)
medianB <- round(problem_6_CompB$median,3)
sdB <- round(problem_6_CompB$sd,3)
sampleB <- problem_6_CompB$n

problem_6_CompC <- favstats(Mystery$C)
meanC <- round(problem_6_CompC$mean,3)
medianC <- round(problem_6_CompC$median,3)
sdC <- round(problem_6_CompC$sd,3)
sampleC <- problem_6_CompC$n

problem_6_CompD <- favstats(Mystery$D)
meanD <- round(problem_6_CompD$mean,3)
medianD <- round(problem_6_CompD$median,3)
sdD <- round(problem_6_CompD$sd,3)
sampleD <- problem_6_CompD$n

problem_6_CompE <- favstats(Mystery$E)
meanE <- round(problem_6_CompE$mean,3)
medianE <- round(problem_6_CompE$median,3)
sdE <- round(problem_6_CompE$sd,3)
sampleE <- problem_6_CompE$n

```

6. If you were saving for retirement, which of these five companies would be the most attractive investment?  Justify your answer.
**Compared to your answer for this question in lesson 3 (question 5), did it make a difference to compare the spread of the data when making a decision for this problem **

```{r,include=FALSE}
#1st column in the bird flu table
Observation = c(39.1,39.5,38.9,39.2,39.9,39.7,39.0)
birdFlu = data.frame(Observation = c(39.1,39.5,38.9,39.2,39.9,39.7,39.0))
#2nd column
birdFlu$Deviation = round(birdFlu$Observation - mean(birdFlu$Observation),3)
#3rd Column
birdFlu$Squared = round((birdFlu$Observation - mean(birdFlu$Observation))^2,3)
#last row (Mean =## Sum =0 and Sum = ##)
##Problem with this: The middle column does NOT add up to zero! Checked it on excel as well. Doesn't add to zero 
Rows = nrow(birdFlu)
birdFlu[Rows + 1,] = c(paste("Mean = ",round(mean(birdFlu$Observation,3)),sep = ""),paste("Sum = ",round(sum(birdFlu$Deviation),3),sep = ""),paste("Sum = ",round(sum(birdFlu$Squared),3),sep = ""))

colnames(birdFlu) = c("Observation, x","Deviation from the mean","Squared Deviation from the Mean")

Variance = var(Observation)
SD = sd(Observation)
```


7.	Create a table to compute the standard deviation "by hand" using the data Severe Bird Flu Data, available on the wiki page, [(Lesson 03 Describing Quantitative Data; Shape, Center and Spread)](http://statistics.byuimath.com/index.php?title=Lesson_3:_Describing_Quantitative_Data:_Shape,_Center_%26_Spread). 

```{r, include = FALSE}
MysteryMelt = melt(Mystery, na.rm = TRUE)
MysteryPlotReady = subset(MysteryMelt, variable != "Year")

Hist_L3_Prob7A <- ggplot(MysteryPlotReady, aes(x = value)) + geom_histogram(binwidth = 15,colour="white") + facet_wrap(~variable,nrow = 5)

ggsave(Hist_L3_Prob7A,filename="../images/Hist_L3_Q07.png",width=4,height=4)
```


```{r,include=FALSE}
if(params$key==TRUE){
#solutions

  Solution01 = data.frame(Part=c("Shape","Center","Spread"),
                          Solution = c("c. Symmetric, bell-shaped, skewed right, and skewed left, uniform, unimodal, bimodal, or multimodal",
                                       "a. Average, mean, median, mode",
                                       "b. Standard deviation, variance"))
  
  Solution02 = data.frame(Part="-",Solution = "The standard deviation is a measure of the spread in a distribution.  If the standard deviation is small then the data tend to be closer together.  If the standard deviation is large, the data is usually more spread out.")
  
  Solution03 = data.frame(Part="-",Solution = "Answers will vary")
  
  Solution04 = data.frame(Part="-",Solution = "Sample variance is also a measure of spread in a data set.  The sample variance is an estimator of the true population variance.  The sample variance is just the sample standard deviation squared or $s^2$.")
  
  Solution05 = data.frame(Part=c("Percentiles","Quartiles","Five-number summary","Boxplot"),
                          Solution=c("A percentile is a number such that a specified percentage of the data are at or below this number, such as at the 50th percentile, this value is as large or larger than 50% of all the data.",
                                       "Quartiles are just three special percentiles, $25^{th}$, $50^{th}$, and $75^{th}$ percentiles divide the data into fourths.",
                                       "The five number summary is comprised of the minimum, first quartile, median, third quartile, and maximum.  It is just a good way to summarize the data into just five numbers.",
                                       "A boxplot is a graphical representation of the five-number summary."))
  
  Solution06 = data.frame(Part = "-",Solution = "Answers may vary, students should choose a company and justify why. They will probably choose B or C, because they have the highest means.")
  
  #Solution10 = data.frame(Part = "-",Solution = )
  
} # end params == TRUE.  We use this so it doesn't have to run for non answer key path. 
```


```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){
  cat("-->")
   cat("\n\n## Solutions\n\n")
      cat("\n\n **Please note that the steps show rounded numbers, but that the final answers to the problems are calculated without rounding.**")
}
```



```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){

    all_solutions = sort(ls(pattern="Solution"))
    key_list = NULL
    for (i in 1:length(all_solutions)){
      temp = get(all_solutions[i])
      temp$Solution = as.character(temp$Solution)
      key_list = rbind(key_list,data.frame(Problem=i,temp))
    }
    
      pander(key_list,split.cell = 80, split.table = Inf,justify = c( 'center', 'left',"left"))
} # end params == TRUE.  We use this so it doesn't have to run for non answer key path.

```

