---
title: 'Lesson 19: Inference for Several Proportions, Goodness-of-Fit, and Independence'
author: "Preparation"
output:
  pdf_document: default
  html_document:
    css: buttondown.css
  word_document: default
params:
  key: no
  plinks: no
  keyname: "blank"
  docname: "blank"
md_document:
  variant: markdown_phpextra
---

```{r,results='asis',echo=FALSE}
if(params$key==TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$keyname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$keyname,"pdf",sep="."),")\n")
  cat("<!--")
  cat("\n")
} 
if(params$key!=TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$docname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$docname,"pdf",sep="."),")\n")
  cat("\n")
}
```

**Directions: Please fill in Part I as you study the Reading Assignment. Once you finish the reading, complete the questions on Part II. You may use your notes, the key, and the help videos. Be sure to take this completed assignment to your group meeting where you can ask and help answer questions on this assignment.**

```{r,include=FALSE}
library(ggplot2)
library(ggthemes)
library(pander)
panderOptions('keep.line.breaks',TRUE)
source("../scripts/ggQQline.R")
source("../scripts/normTail.R")
source("../scripts/221_Interactive_Functions.R")
```

## Problems

**Part I:**  Use the information in the reading assignment to complete these questions.

1.	For a Chi-Square Test for Independence, what are the null and alternative hypotheses?

2.	True/False: If the null hypothesis is rejected, then the row variable and column variable are not independent. 


**Part II:**  

Use the following data set to answer question 3.

```{r,include=FALSE}
inFavorRepublican = 250
AgainstRepublican = 50

inFavorDemocrat = 120
AgainstDemocrat = 80

confLevel = 95
alpha = 1 - (confLevel/100)

Prop8 = cbind(Republican = c(inFavorRepublican,AgainstRepublican), Democrat = c(inFavorDemocrat,AgainstDemocrat))
rownames(Prop8) = c("In Favor of Prop 8","Against Prop 8")

results = chisq.test(Prop8,correct = FALSE)
reqMet = reqMetChi(results$expected)

df = results$parameter

chi = results$statistic

pvalue = results$p.value
```

```{r, echo=FALSE, results='asis'}
pander(data.frame(Population=c("In Favor of Prop 8", "Against Prop 8"), Republican=c(inFavorRepublican,AgainstRepublican),Democrat=c(inFavorDemocrat,AgainstDemocrat)))
```

3.	Ambitious Intro to Stats students wanted to conduct a survey among church members and asked them about their party affiliation and whether or not they were in favor of Proposition 8.  They would like to see if there is an association between party affiliation and voting preference for Prop 8 with a level of significance of $\alpha = `r alpha`$. 

    a. State the null and alternative hypotheses.

    b. What are the requirements for this test?

    c. Are the requirements met?

    d. Calculate the $\chi^2$ (Chi-Square) statistic?

    e. Determine the Degrees of Freedom for this Test?

    f. What is the P-value of the $\chi^2$ statistic?

    g. Do you reject the null hypothesis or fail to reject the null hypothesis?

    h. State your conclusions based on your p-value and based on a level of significant of $\alpha=`r alpha`5$. 

```{r,include=FALSE}
if(params$key==TRUE){
#solutions

  Solution01=data.frame(Part = "-",Solution = "$H_0: \\text{ The row variable and the column variable are independent}$ \\\n $H_a: \\text{  The row variable and the column variable are not independent}$")
  
  Solution02=data.frame(Part = "-",Solution = "True")
  
  Solution03=data.frame(Part = LETTERS[1:8],Solution = c("$H_0: \\text{ Party affiliation and whether they were in favor or against Prop 8 are independent}$ \\\n $H_a: \\text{ Party affiliation and whether they were in favor or against Prop 8 are not independent}$",
                                                         "That all expected cell counts are greater than 5",
                                                         reqMet,
                                                         paste("$\\chi^2= ",round(chi,2),"$",sep = ""),
                                                         paste("$df = ",df,"$",sep = ""),
                                                         paste("$\\text{P-value} = ",round(pvalue,3),"$",sep = ""),
                                                         failOrNot(pvalue,alpha),
                                                         paste("We have ",sufficientOrNot(failOrNot(pvalue,alpha))," evidence to show that the party affiliation and the voters' views on Proposition 8 are not independent.")))
  
} # end params == TRUE.  We use this so it doesn't have to run for non answer key path. 
```


```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){
  cat("-->")
   cat("\n\n## Solutions\n\n")
      cat("\n\n **Please note that the steps show rounded numbers, but that the final answers to the problems are calculated without rounding.**")
}
if(params$key==TRUE) cat("-->")
  if(params$key==TRUE) cat("## Solutions\n\n")
      if(params$key==TRUE) cat("\n\n")
```



```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){

    all_solutions = sort(ls(pattern="Solution"))
    key_list = NULL
    for (i in 1:length(all_solutions)){
      temp = get(all_solutions[i])
      temp$Solution = as.character(temp$Solution)
      key_list = rbind(key_list,data.frame(Problem=i,temp))
    }
    
      pander(key_list,split.cell = 80, split.table = Inf,justify = c( 'center', 'left',"left"))
} # end params == TRUE.  We use this so it doesn't have to run for non answer key path.

```
