---
title: 'Lesson 2: The Statistical Process & Design of Studies'
author: "Preparation"
output:
  html_document:
    css: buttondown.css
  pdf_document: default
  word_document: default
params:
  key: yes
  plinks: no
  keyname: "blank"
  docname: "blank"
md_document:
  variant: markdown_phpextra
---

```{r,results='asis',echo=FALSE}
if(params$key==TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$keyname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$keyname,"pdf",sep="."),")\n")
  cat("<!--")
  cat("\n")
} 
if(params$key!=TRUE){
  if(params$plinks==TRUE) cat("* [Word Document](",paste(params$docname,"docx",sep="."),")\n")
  if(params$plinks==TRUE) cat("* [PDF Document](",paste(params$docname,"pdf",sep="."),")\n")
  cat("\n")
}
```

**Directions: Please fill in Part I as you study the Reading Assignment. Once you finish the reading, complete the questions on Part II. You may use your notes, the key, and the help videos. Be sure to take this completed assignment to your group meeting where you can ask and help answer questions on this assignment.**

```{r,include=FALSE}
library(ggplot2)
library(ggthemes)
library(pander)
panderOptions('keep.line.breaks',TRUE)
source("../../scripts/ggQQline.R")
source("../../scripts/normTail.R")
source("../../scripts/221_Interactive_Functions.R")
```

## Problems

**Part I: Use the information in the reading assignment to complete these questions.**

1. State the five steps of the statistical process and provide a brief description for each step.

2. Match the following statistical terms with their appropriate definitions.    
    ___ Simple Random Sample (SRS)    
    ___ Stratified Sampling   
    ___ Systematic Sampling   
    ___ Cluster Sampling    

    a. Subjects are in groups according to similarity of some characteristics (e.g. age, income level, political party); specifically take a simple random sample from each group. (Homogeneous)
    
    b. Subjects are in "some" sequential order. Randomly select one subject (to start your sampling) then ask every $k^{th}$ subject.
    
    c. Computer generated; draw from a hat, etc.
    
    d. Subjects are put into groups that "hopefully" represent the population. Randomly select one or more groups and sample everybody in that group. (Heterogenous)

3. Provide a brief description for each of the follow terms.    
    
    Experiment -
    
    Observational Study -
    
    Treatment -
    
    Response Variable -
    
    Subjects -

4. Define the Null and Alternative Hypothesis.

5. Define the P-value.

**Part II:** 

```{r,include=FALSE}
# N = sample size, Kth = "every Kth person"

bN = 2000000
bKth = 1000

cN = 1059

dN = 200

eHospitals = 30

```


6. Determine the correct sampling method for each scenario.

    a. A medical student researches the prevalence of aspirin usage by questioning all of the patients who enter her clinic for treatment.
    
    b. The Colorado Department of Health obtains an alphabetical list of `r formatC(bN, big.mark = ",",format = "fg")` adult Colorado residents and constructs a pool of survey subjects by selecting every `r formatC(bKth, big.mark = ",",format = "fg")`th name on the list after randomly choosing one person from the first `r formatC(bKth, big.mark = ",",format = "fg")` on the list.
    
    c. In a Gallup Poll of `r formatC(cN, big.mark = ",",format = "fg")` adults, the interview subjects were selected by using a computer to randomly generate telephone numbers that were then called.
    
    d. A dietitian has partitioned people into age categories of under 18, 18-49, 50-69, and over 69.  She is surveying a simple random sample of `r formatC(dN, big.mark = ",",format = "fg")` members from each category.
    
    e. A Johns Hopkins University researcher surveys all cardiac patients in each of `r formatC(eHospitals, big.mark = ",",format = "fg")` randomly selected hospitals.
    
```{r,include=FALSE}
bN = 100
```


7. The American Medical Association wanted to determine general patient satisfaction with the patient's primary care physician. Determine the type of sampling method used in the following scenarios.

    a. They take a simple random sample of doctors and then sample all of the patients that the doctor works with.

    b. They put all of the residents' names in a computer and use a random number generator to get a sample of `r formatC(bN, big.mark = ",",format = "fg")` patients.

    c. They divide the residents up by gender and then do a simple random sample of residents within each gender.

    d. They randomly pick key intersections in the city and then picking every fifth person that walks by after randomly picking one of the first five people to be in the sample.

```{r,include=FALSE}
aMenN = 300
aWomenN = 300

bTotalN = 1000
```


8. Determine for each scenario whether it is an observational study or an experimental design. If it is an experimental design, determine the response, treatment and subjects.

    a. The Center for Disease Control obtains current flu data by randomly polling `r formatC(aMenN, big.mark = ",",format = "fg")` men and `r formatC(aWomenN, big.mark = ",",format = "fg")` women.

    b. You plan to conduct an experiment to test the effectiveness of Sleepeze, a new drug that is supposed to reduce insomnia, so you randomly divide `r formatC(bTotalN, big.mark = ",",format = "fg")` patients into two groups, where half get Sleepeze and the other a placebo.  You record the gender of each subject.  You will then observe how many hours of sleep they get in a week.

```{r,include=FALSE}
TotalN = 224

Group1 = 0.5
Group2 = 1
```


9. In September 2004, Nucryst Pharmaceuticals, Inc. announced the results of its first human trial of NPI 32101, a topical form of its skin ointment.  A total of `r formatC(TotalN, big.mark = ",",format = "fg")` patients diagnosed with skin irritations were randomly divided into three groups as part of a double-blind, placebo-controlled study to test the effectiveness of the new topical cream.  The first group received a `r Group1`% cream, the second group received a `r Group2`% cream, and the third group received a placebo. Groups were treated twice daily for a 6-week period and the degree of skin irritation was recorded at that time.

    a. Is this a designed experiment or an observational study?

    b. What are the subjects?

    c. What is the response variable in the experiment?

    d. What are the treatments?


```{r,include=FALSE}
if(params$key==TRUE){
  #solutions

   Solution01 = data.frame(Part = LETTERS[1:5], 
                          Solution = c("Design the Study - State a research question, what needs to be done to answer the research question.  What is the population? What kind of data needs to be collected?",
                                       "Collect Data - How is the sample collected, and going out to actually obtain the data?",
                                       "Describe the Data - Creating graphs or calculating statistics to help visualize and describe the data.",
                                       "Make Inferences - Using the information contained in a sample to draw conclusions about a population.",
                                       "Take Action - Determine which action to take based on the results of the study."))
   
   Solution02 = data.frame(Part = LETTERS[1:4],
                           Solution = c("c- Simple Random Sample (SRS)",
                                        "a - Stratified Sampling",
                                        "b - Systematic Sampling",
                                        "d - Cluster Sampling"))
   
   Solution03 = data.frame(Part = "-",
                           Solution = "Experiment - Researchers control the conditions under which measurements are made.\\\n \\\n Observational Study - Researchers simply observe what happens without controlling the conditions under which measurements are made. \\\n \\\n Treatment - The new or experimental condition that is imposed on the subjects. \\\n \\\n Response Variable - The variable that changes or responds to the treatment. \\\n \\\n Subjects - Participants in a study ")
   
   Solution04 = data.frame(Part = "-",
                           Solution = "The Null Hypothesis ($H_0$) is the foundational assumption about the population, the 'status quo' \\\n \\\n The Alternative Hypothesis ($H_a$) is the statement of what is true if the null hypothesis is rejected.  It is also a statement about the population.")
   
   Solution05 = data.frame(Part = "-",Solution="P-Value: The probability that a result will be observed that is as extreme or more extreme than the information observed, assuming that that the null hypothesis is true.")
   
   Solution06 = data.frame(Part = LETTERS[1:5],
                          Solution = c("Convenience Sampling",
                                       "Systematic Sampling",
                                       "Simple Random Sampling",
                                       "Stratified Sampling",
                                       "Cluster Sampling"))
   
   Solution07 = data.frame(Part = LETTERS[1:4],
                           Solution = c("Cluster Sampling",
                                        "Simple Random Sampling",
                                        "Stratified Sampling",
                                        "Systematic Sampling"))
   
   Solution08 = data.frame(Part = LETTERS[1:2],
                           Solution = c("Observational Study",
                                        paste("Experimental Design \\\n \\\n Response Variable - Hours of sleep they get in a week \\\n \\\n Treatment - Sleepeze or placebo \\\n \\\n Subjects - ",formatC(bTotalN, big.mark = ",",format = "fg")," insomnia patients",sep = "")))
   
   Solution09 = data.frame(Part = LETTERS[1:4], 
                           Solution = c("Designed Experiment",
                                        paste(formatC(TotalN,big.mark = ",",format = "fg")," patients diagnosed with skin irritations",sep = ""),
                                        "The degree of skin irritations observed in the patients",
                                        paste(Group1,"% cream, ",Group2,"% cream, and placebo",sep = "")))
   
} # end params == TRUE.  We use this so it doesn't have to run for non answer key path. 
```

```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){
  cat("-->")
   cat("\n\n## Solutions\n\n")
      cat("\n\n **Please note that the steps show rounded numbers, but that the final answers to the problems are calculated without rounding.**")
}
```



```{r,echo=FALSE,results='asis'}
if(params$key==TRUE){

    all_solutions = sort(ls(pattern="Solution"))
    key_list = NULL
    for (i in 1:length(all_solutions)){
      temp = get(all_solutions[i])
      temp$Solution = as.character(temp$Solution)
      key_list = rbind(key_list,data.frame(Problem=i,temp))
    }
    
      pander(key_list,split.cell = 80, split.table = Inf,justify = c( 'center', 'left',"left"))
} # end params == TRUE.  We use this so it doesn't have to run for non answer key path.

```